Cycles of the logistic map

### 2013-06-03 doc/pc-ijbc.tex
add lsfit.ma, T14 polynomials

\begin{abstract}
The onset and bifurcation points of the $n$-cycles of
  a polynomial map are located
  through a characteristic equation
  connecting the cyclic polynomials of the cycle points.
The polynomials satisfied by the parameters
  of the logistic, H\'enon, and cubic maps
  at the onset and bifurcation points are obtained
  for $n$ up to 13, 9, and 8,
  respectively.
\end{abstract}

@@

Consider the logistic map \cite{may, strogatz}:
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv r \, x_k \, ( 1 - x_k ).
\label{eq:logmap}
\end{equation}
%
%The iterated sequence $x_1$,
%  $x_2 = f(x_1)$,
%  $x_3 = f(x_2)$,
%  $\ldots$,
%can be conveniently visualized on the cobweb plot,
%  as shown in \reffig{cobweb}.
%%
%Starting from $(x_1, x_1)$ on the diagonal,
%  each vertical arrow takes $(x_k, x_k)$ to $(x_k, y)$,
%  where $y = x_{k+1} = f(x_k)$;
%by the following horizontal arrow,
%  it returns to the diagonal at
%  $(y, y) = (x_{k+1}, x_{k+1})$,
%  which starts the next iteration.
%
%
%\begin{figure}[h]
%  \begin{center}
%  \begin{minipage}{\linewidth}
%        \includegraphics[angle=-90, width=\linewidth]{cobweb.pdf}
%  \end{minipage}%
%  \end{center}
%  \caption{\label{fig:cobweb}
%  Cobweb plot of the logistic map.}
%\end{figure}
%
%
After many iterations, there can be three ultimate patterns:
  (i) a \emph{fixed point} (including infinity),
      %e.g., \reffig{cobweb}(a);
  (ii) a periodic \emph{cycle},
      %e.g., \reffigs{cobweb}(b)-(e);
or
  (iii) a chaotic trajectory.
      %e.g., \reffig{cobweb}(f).
We only discuss the first two cases here.

A fixed point is a solution of $x^* = f(x^*)$.
%
It is stable if a small deviation from $x^*$
  gets reduced by successive iterations.
%
For a differentiable $f$,
  this is equivalent to $|f'(x^*)| \le 1$
  \cite{strogatz}.
%
An $n$-cycle is a sequence $x_1, \dots, x_n$,
  such that $n$ is the smallest positive integer
  that satisfies $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
Obviously, any $x_k$ in an $n$-cycle of $f$ is a fixed point of $f^n$.
%
%\big[but the converse is untrue, for a fixed point of $f^n$
%  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
%  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big].
%
%
Thus, a cycle can be classified as stable or unstable
  according to the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_1) \cdots f'(x_n) \Big| \le 1,
\label{eq:der}
\end{equation}
%
%
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz}.
%
%
%
The cycle and its stability depend on the parameter $r$,
%
and a stable cycle corresponds to a window
  between the onset point $r_a$ and bifurcation point $r_b$.
%
There are generally multiple such windows for a single $n$,
  but all $r_a$ are zeros of a polynomial,
  and all $r_b$ another.
%
Our goal is thus to find the two polynomials for a given $n$.
%
The problem has been tackled by several previous studies
  \cite{brown1, brown2, stephenson1, stephenson2, stephenson3,
  saha, bechhoefer, gordon, burm, zhang, bailey1, bailey2, kk1, lewis}.
Below we present a simple algorithm to do so.
%
%
%
%The window boundaries can be classified
%  to onset and bifurcation points according to \refeq{der};
%Here, if $r > 0$, then
%  $r_a$ ($r_b$) are the onset (bifurcation) points.
%

@@

\subsection{Strategy}

Since the $n$-cycles points constitute
  a subset of the fixed points of $f^n$,
we will first find
  the polynomials at the stability boundaries
  for the fixed points of $f^n$
  (\refsecs{cyclic}-\ref{sec:examples}).
The polynomials contain
  factors contributed by short $d$-cycles with $d|n$,
  which are also fixed points of $f^n$.
The factors are removed in \refsec{primfac}.

At first glance, the polynomials of the fixed points of $f^n$
  can be obtained by brute force.
We can solve
  \refeqs{logmaps} and express $x_1, \ldots, x_n$
  in terms of $R$,
  and then plug the solution into \refeq{der}.
The result would contain $R$ only,
  and therefore be the answer.
But since \refeqs{logmaps} are nonlinear,
  it is nontrivial to reduce the final equation of $R$
  into a polynomial one.
Nonetheless, on a computer, one can
  construct a polynomial Gr\"obner basis \cite{kk1},
  which gradually eliminates $x_k$ out of the $n+1$ equations.
If $x_2, \dots, x_n$ are eliminated first,
  there are only two equations left:
  $f^n(x_1) = x_1$
  and
  $\frac{d}{dx}f^n (x_1) = \pm 1$
  [\refeq{der}];
  the final elimination of $x_1$ can then be accomplished
  by setting the resultant of the two equations
  to zero \cite{burm},
  leading to the desired polynomial equation of $R$.
%
These approaches, however, do not
  exploit the cyclic structure of \refeqs{logmaps},
  and can thus be improved by the following alternative.

@@

For $n = 2$, we have %three square-free cyclic polynomials
$\{C_0, C_1, C_{12}\}
  = \{1,  x_1 + x_2, x_1 x_2 \}$
and
$\Lambda = 4 \, x_1 x_2$.
So
$\Lambda \, C_0 = 4 \, C_{12}$,
$\Lambda \, C_1 %= 4 \, {x_1}^2 \, x_2 + 4 \, x_1 \, {x_2}^2
  = -8 \, R \, C_0 + 4 \, (R + 1) \, C_1$
by the example after \refthm{sqrfree},
$\Lambda \, C_{12}
= 4 \, (R - x_2) (R - x_1) = 4 \, (R^2C_0-RC_1+C_{12})$,
and \refeq{xr} reads
%\[
%\Lambda
%  \left( \begin{array}{c}
%  C_0 \\
%  C_1 \\
%  C_{12}
%  \end{array} \right)
% =
%  \left( \begin{array}{ccc}
%  0           & 0         & 4 \\
%  -8R         & 4(R+1)    & 0 \\
%  4R^2        & -4R       & 4
%  \end{array} \right)
%  \left( \begin{array}{c}
%  C_0 \\
%  C_1 \\
%  C_{12}
%  \end{array} \right).
%\]
%
%
%
%The characteristic equation (\ref{eq:xr}) reads
%
%
%
\begin{align}
A_2 \equiv (4 - 4R - \lambda) [(4 R-\lambda)^2 - 4 \lambda] = 0.
  \tag{\ref{eq:xr}-2}
\label{eq:xr2}
\end{align}
%
%
%
Only the first factor, denoted as $P_2\Rlam = 4-4R -\lambda$,
  corresponds to the 2-cycle  (cf. \refsec{primfac}).
%  while the second factor is from the fixed points
So $R = 1 - \lambda/4$,
and the onset ($\lambda = +1$) value is $R_a = 3/4$ %($r_a = 3$)
and the bifurcation ($\lambda = -1$) value is $R_b = 5/4$. %($r_b = 1+\sqrt 6$).
%
%Note that the onset point of the 2-cycle
%  is located at $R = 3/4$,
%  where the fixed point bifurcates \cite{strogatz}.
%  \big[compare \reffigs{cobweb}(a) and (b)\big].

@@

For the $n = 4$ case,
%\cite{brown1, stephenson1, stephenson3, bailey1, bailey2, kk1, lewis},
%$C_0 = 1$,
%$C_1 = x_1 + \cdots + x_4$,
%$C_{12} = x_1 x_2 + \cdots + x_4 x_1$,
%$C_{13} = x_1 x_3 + x_2 x_4$,
%$C_{123} = x_1 x_2 x_3 + \cdots + x_4 x_1 x_2$,
%$C_{1234} = x_1 x_2 x_3 x_4$,
%and
%$\Lambda = 16 \, C_{1234}$.
%
we only list the final \refeq{xr}: %characteristic equation (\ref{eq:xr}):
%after the square-free reduction:
%(with six square-free cyclic polynomials):
%
%\begin{small}
\begin{align}
A_4
%& = \left|
%\begin{array}{cccccc}
% \lambda  & 0               & 0                 & 0     & 0     & -16 \\
% 64 R     & \lambda-16(R+1) & 16 R              & 0     & -16R  & 0 \\
% -64 R^2  & 16R(R+2)        & \lambda-16(R^2+1) & -32 R & 16R   & 0 \\
% -32 R^2  & 16R(R+1)        & -16R              & \lambda-16(R^2+1)  & 0 & 0 \\
% 64 R^3   & -16R^2(R+3)     & 16R(R+2)        & 32 R (R+1)       & \lambda-16(R+1) & 0 \\
% -16R^4   & 16R^3           & - 16R^2           & -16R^2            & 16R & \lambda - 16
%\end{array}
%\right | \nonumber  \\
& \equiv
    \big[
    4096 R^6-12288 R^5 +
    256 (\lambda + 48) (R^4- R^3) \notag \\
& \quad
    -16 (\lambda + 32)(\lambda - 16) R^2
      -(\lambda - 16)^3 \big] \notag \\
& \quad
    \big[(16 R^2 + \lambda)^2 - 16 (2R+1)^2 \lambda \big] \notag \\
& \quad
    \big[ 16 (R-1)^2 - \lambda \big]
 = 0.
\tag{\ref{eq:xr}-4}
\label{eq:xr4}
\end{align}
%\end{small}
%
%
%
The first (sextic) factor, denoted as $P_4\Rlam$,
  is responsible for the 4-cycle.
%
At the onset point, % $\lambda = 1$.
we have %it can be factorized as
$ P_4(R, 1)
= (
    5 - 4R
  )
\, \big[
  (4R + 1)^2 + 4
  \big]
\, \big[
  108 - (4R - 3)^3
  \big]$,
%
%
%
with two real roots:
$R_a' = 5/4$
%\big($r_a' = 1+\sqrt{6} \approx 3.4495$\big)
for the period-doubled 2-cycle,
  %\big[compare \reffigs{cobweb}(b) and (d)\big],
and
$R_a = (3+\sqrt[3]{108})/4$
%\big($r_a = 1+\sqrt{4+\sqrt[3]{108}} \approx 3.9601$\big)
for an original 4-cycle. %\big[\reffig{cobweb}(e)\big].
%
At the bifurcation point, % ($\lambda = -1$),
we have $P_4(R, -1)
 = 4096 R^6 - 12288 R^5 + 12032 (R^4 - R^3)
  + 8432 R^2 + 4913$.
%which upon $R \rightarrow r(r-2)/4$ yields the same polynomial
%obtained previously
%\cite{stephenson1, stephenson3, bailey2, kk1, lewis}.
%
%The only two positive roots $r_b \approx 3.9608$
%and $r_b' \approx 3.5441$ of the sextic equation
%correspond to $r_a$ and $r_a'$ respectively.
%
%Since the 4-cycle problem were tackled mainly on computers
%  in the previous studies,
%  we also supply a compact elementary derivation
%  in the Appendix.

\subsection{\label{sec:primfac}$n$-cycle polynomial}

@@

%\begin{theorem}
%  The polynomial $P_n\Rlam$ of all $n$-cycles
%  is a factor of the characteristic polynomial $A_n\Rlam$,
%  and can be computed as
  \begin{equation}
    P_n\Rlam
    = \prod_{cd = n} B_d^c\Rlam^{\mu(c)},
  \label{eq:primfac}
  \end{equation}
where
  $B_d^c\Rlam \equiv \prod_{k=1}^c A_{d}(R, e^{2k\pi i/c} \lambda^{1/c})$,
  and
  $\lambda^{1/c}$ is a complex $c$th root of $\lambda$;
  the outer product runs over all divisors $d$ of $n$.
%\end{theorem}
%
%
%\begin{remark}
%Here, ``minimal'' means that the polynomial contains all zeros.
%It happens to be irreducible for a general $\lambda$.
%At $\lambda = 1$, the polynomial of $R$, however, can be factorized (see later).
%\end{remark}
%
%\begin{remark}
The M\"obius function $\mu(n)$ is $(-1)^k$
  if $n$ is the product of $k$ distinct primes,
  or 0 if $n$ is divisible by a square of a prime;
%Thus,
$\mu(n) = 1, -1, -1, 0, -1, 1, -1, \ldots,$
  from $n = 1$
%The $\mu(n)$ is useful for inversion:
%$g(n) = \sum_{d|n} \mu(n/d) \, h(d)$
%if and only if $h(n) = \sum_{d|n} g(d)$
 \cite{hardy}.
%\end{remark}
%

%\begin{remark}
%$B_{d,c}\lambda$ is a polynomial of $\lambda$.
Despite the radical argument $\lambda^{1/c}$,
  the product
  $B_d^c\Rlam = \prod_{k=1}^c A_d(R, e^{2k\pi i/c} \lambda^{1/c})$
  is a polynomial of $\lambda$,
  for it is invariant under
  $\lambda \rightarrow e^{2\pi i} \lambda$.
For convenience, we will drop the $R$ in ``$\Rlam$'' below till the end the next section.
Particularly, $B_n^1\lam = A_n\lam$.
We also have $\deg_\lambda B_d^c\lam = \deg_\lambda A_d\lam$.
%
%\end{remark}

@@

%\subsection{\label{sec:degR}Degrees in $R$}
%
%
%
%In the numerical implementation
%  (cf. \refsec{end}),
%  the degrees in $R$ of the polynomials are needed
%  for evaluating the determinant by Lagrange interpolation.
%We now show that
%\begin{equation}
%\deg_R A_n\Rlam = \textstyle\sum_{d|n} \varphi(n/d) 2^{d-1},
%\label{eq:degRA}
%\end{equation}
%%
%The first few values are
%1, 3, 6, 12, 20, 42, 70, 144, 270, 540, \dots, from $n =1$.%1034, 2112, 4108, \dots,
%
%From \refeq{xr},
%  every term in the determinant $|\vct A|$
%  can be written as
%  $\prod_{p} A_{p,q}$,
%  in which, $p$ runs through all cyclic polynomials,
%  and $\{q = q(p)\}$ is a permutation of $\{p\}$.
%Thus, $\deg_R |\vct A|$ is
%  reached by the term that maximizes the product.
%Recall that $A_{p,q}$ results from reducing
%  $\Lambda(\vx{n}) \, C_p(\vx{n})$ to $C_q(\vx{n})$
%  by a sequence of substitutions ${x_k}^2 = R-x_{k+1}$.
%It follows that to increase $\deg_R$ by 1,
%  $\deg_x$ is decreased by at least 2.
%Accordingly,
%  $\deg_R A_{p,q} \le \deg_x(\Lambda \, C_p) - \deg_x C_q$.
%Now running $p$ and $q$ through the $\NB$ cyclic polynomials
%  and using $\deg_x \Lambda(\vx{n}) = n$
%  yields
%\[
%  \deg_R |\vct A | = \deg_R \textstyle\prod_{p} A_{p,q} \le n \, \NB(n).
%\]
%%
%The equality is achieved, if each $q$ is the complement $\bar p$ of $p$,
%  that is, if $C_p$ is generated from $\prod_k x_k^{e_k}$,
%  $C_{\bar p}$ can be generated from $\prod_k x_k^{1-e_k}$
%%  (which may be the same polynomial as $C_p$).
%For example, for $n=4$,
%  $C_1 = x_1 + \cdots + x_4$
%  is the complement of
%  $C_{123} = x_2 x_3 x_4 + \cdots + x_1 x_2 x_3$,
%  while $C_{12}$ is its own complement.
%Finally, we reach \refeq{degRA} by \refeq{necklace}.
%
%
%Now, taking the degrees in $R$ of \refeq{primfac} and \refeq{origfac}
%  yields two known formulas
%\cite{mira, stephenson1, blackhurst}:
%%
%\begin{subequations}
%\begin{align*}
%\deg_R P_n\Rlam &\equiv \beta(n)
%  = \sum_{d|n} \mu\Big(\frac{n}{d}\Big) 2^{d-1}, \\
%\deg S_n(R) &= \beta(n)
%  - \sum_{d | n, d < n} \beta(d) \, \varphi\Big(\frac{n}{d}\Big).
%\end{align*}
%\label{eq:degR}
%\end{subequations}
%%
%The first few $\deg_R P_n\Rlam$ are
%  1, 1, 3, 6, 15, 27, 63, 120, 252, 495, \dots, %1023, 2010, 4095, \dots,
%and $\deg S_n(R) = $
%  1, 0, 1, 3, 11, 20, 57, 108, 240, 472, \dots, % 1013, 1959, 4083, \dots,
%from $n = 1$.
%

@@

The most time-consuming step of the algorithm in \refsec{algo}
is the evaluation of the determinant $A_n\Rlam$ in the characteristic
equation (\ref{eq:xr}).
%
When implemented on a computer,
it is helpful to use Lagrange interpolation:
we evaluate $A_n\Rlam$ at $\deg_R A_n + 1$ values of $R$,
e.g., $R = 0, \pm1, \pm2,\ldots$, then piece them together
to a polynomial.
%
By using \refeq{primfac}, we only need $\deg_R P_n + 1$ values
  to directly compute $P_n\Rlam$.
%
%
Here,
$\deg_R A_n = \sum_{d|n} \varphi(n/d) 2^{d-1}$,
and $\deg_R P_n = \sum_{d|n} \mu(n/d) 2^{d-1}$
  \cite{mira, stephenson1}.
%
The strategy also allows a trivial parallelization.

>>>>

\begin{abstract}
The onset and bifurcation points of the $n$-cycles of
  a polynomial map are located
  through a characteristic equation
  connecting the cyclic polynomials of the cycle points.
The polynomials satisfied by the parameters
  of the logistic, H\'enon, and cubic maps
  at the onset and bifurcation points are obtained
  for $n$ up to 14, 9, and 8,
  respectively.
\end{abstract}

@@

Consider the logistic map \cite{may, strogatz}:
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv r \, x_k \, ( 1 - x_k ).
\label{eq:logmap}
\end{equation}
%
%The iterated sequence $x_1$,
%  $x_2 = f(x_1)$,
%  $x_3 = f(x_2)$,
%  $\ldots$,
%can be conveniently visualized on the cobweb plot,
%  as shown in \reffig{cobweb}.
%%
%Starting from $(x_1, x_1)$ on the diagonal,
%  each vertical arrow takes $(x_k, x_k)$ to $(x_k, y)$,
%  where $y = x_{k+1} = f(x_k)$;
%by the following horizontal arrow,
%  it returns to the diagonal at
%  $(y, y) = (x_{k+1}, x_{k+1})$,
%  which starts the next iteration.
%
%
%\begin{figure}[h]
%  \begin{center}
%  \begin{minipage}{\linewidth}
%        \includegraphics[angle=-90, width=\linewidth]{cobweb.pdf}
%  \end{minipage}%
%  \end{center}
%  \caption{\label{fig:cobweb}
%  Cobweb plot of the logistic map.}
%\end{figure}
%
%
After many iterations, there can be three ultimate patterns:
  (i) a \emph{fixed point} (including infinity),
      %e.g., \reffig{cobweb}(a);
  (ii) a periodic \emph{cycle},
      %e.g., \reffigs{cobweb}(b)-(e);
or
  (iii) a chaotic or aperiodic trajectory.
      %e.g., \reffig{cobweb}(f).
We only discuss the first two cases here.

A fixed point is a solution of $x^* = f(x^*)$.
%
It is stable if a small deviation from $x^*$
  gets reduced by successive iterations.
%
For a differentiable $f$,
  this is equivalent to $|f'(x^*)| \le 1$
  \cite{strogatz, hao}.
%
An $n$-cycle is a sequence $x_1, \dots, x_n$,
  such that $n$ is the smallest positive integer
  that satisfies $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
Obviously, any $x_k$ in an $n$-cycle of $f$ is a fixed point of $f^n$.
%
%\big[but the converse is untrue, for a fixed point of $f^n$
%  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
%  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big].
%
%
Thus, a cycle can be classified as stable or unstable
  according to the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_1) \cdots f'(x_n) \Big| \le 1,
\label{eq:der}
\end{equation}
%
%
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz, hao}.
%
%
%
The cycle and its stability depend on the parameter $r$,
%
and a stable cycle corresponds to a window
  between the onset point $r_a$ and bifurcation point $r_b$.
%
There are generally multiple such windows for a single $n$,
  but all $r_a$ are zeros of a polynomial,
  and all $r_b$ another.
%
Our aim is thus to find the two polynomials for a given $n$.
%
The problem has been tackled in previous studies
  \cite{brown1, brown2, stephenson1, stephenson2, stephenson3,
  saha, bechhoefer, gordon, burm, zhang, bailey1, bailey2, kk1, lewis}.
Below we present a simple algorithm to do so.
%
%
%
%The window boundaries can be classified
%  to onset and bifurcation points according to \refeq{der};
%Here, if $r > 0$, then
%  $r_a$ ($r_b$) are the onset (bifurcation) points.
%

@@

\subsection{Plan}

Since the $n$-cycles points constitute
  a subset of the fixed points of $f^n$,
we will first find
  the polynomials at the stability boundaries
  for the fixed points of $f^n$
  (\refsecs{cyclic}-\ref{sec:examples}).
The polynomials contain
  extraneous factors contributed by short $d$-cycles with $d|n$,
  which are also fixed points of $f^n$.
The factors are removed in \refsec{primfac}.

At first glance, the polynomials of the fixed points of $f^n$
  can be obtained by brute force.
We can solve
  \refeqs{logmaps} and express $x_1, \ldots, x_n$
  in terms of $R$,
  and then plug the solution into \refeq{der}
  to get a polynomial equation of $R$.
But since \refeqs{logmaps} are nonlinear,
  it is nontrivial to reduce the final equation of $R$
  into a polynomial one.
Nonetheless, on a computer, one can
  construct a polynomial Gr\"obner basis \cite{kk1},
  which gradually eliminates $x_k$ out of the $n+1$ equations.
If $x_2, \dots, x_n$ are eliminated first,
  there are only two polynomial equations left:
  $f^n(x_1) - x_1 = 0$
  and
  $\frac{d}{dx}f^n (x_1) \mp 1 = 0$
  [\refeq{der}];
  the final elimination of $x_1$ can then be accomplished
  by setting the resultant of the two polynomials
  to zero \cite{burm}.
  %leading to the desired polynomial equation of $R$.
%
These approaches, however, do not
  exploit the cyclic structure of \refeqs{logmaps},
  and can thus be improved by the following alternative.

@@

For $n = 2$, we have %three square-free cyclic polynomials
$\{C_0, C_1, C_{12}\}
  = \{1,  x_1 + x_2, x_1 x_2 \}$
and
$\Lambda = 4 \, x_1 x_2$.
So
$\Lambda \, C_0 = 4 \, C_{12}$,
$\Lambda \, C_1 %= 4 \, {x_1}^2 \, x_2 + 4 \, x_1 \, {x_2}^2
  = -8 \, R \, C_0 + 4 \, (R + 1) \, C_1$
by the example after \refthm{sqrfree},
$\Lambda \, C_{12}
= 4 \, (R - x_2) (R - x_1) = 4 \, (R^2C_0-RC_1+C_{12})$,
and \refeq{xr} reads
%\[
%\Lambda
%  \left( \begin{array}{c}
%  C_0 \\
%  C_1 \\
%  C_{12}
%  \end{array} \right)
% =
%  \left( \begin{array}{ccc}
%  0           & 0         & 4 \\
%  -8R         & 4(R+1)    & 0 \\
%  4R^2        & -4R       & 4
%  \end{array} \right)
%  \left( \begin{array}{c}
%  C_0 \\
%  C_1 \\
%  C_{12}
%  \end{array} \right).
%\]
%
%
%
%The characteristic equation (\ref{eq:xr}) reads
%
%
%
{\small
\begin{align}
A_2\Rlam \equiv (4 - 4R - \lambda)
  \left[
    (4 R-\lambda)^2 - 4 \lambda
  \right] = 0.
  \tag{\ref{eq:xr}-2}
\label{eq:xr2}
\end{align}
}
%
%
%
Only the first factor, denoted as $P_2\Rlam = 4-4R -\lambda$,
  corresponds to the 2-cycle  (cf. \refsec{primfac}).
%  while the second factor is from the fixed points
So $R = 1 - \lambda/4$,
and the onset ($\lambda = +1$) value is $R_a = 3/4$ %($r_a = 3$)
and the bifurcation ($\lambda = -1$) value is $R_b = 5/4$. %($r_b = 1+\sqrt 6$).
%
%Note that the onset point of the 2-cycle
%  is located at $R = 3/4$,
%  where the fixed point bifurcates \cite{strogatz}.
%  \big[compare \reffigs{cobweb}(a) and (b)\big].

@@

For the $n = 4$ case,
%\cite{brown1, stephenson1, stephenson3, bailey1, bailey2, kk1, lewis},
%$C_0 = 1$,
%$C_1 = x_1 + \cdots + x_4$,
%$C_{12} = x_1 x_2 + \cdots + x_4 x_1$,
%$C_{13} = x_1 x_3 + x_2 x_4$,
%$C_{123} = x_1 x_2 x_3 + \cdots + x_4 x_1 x_2$,
%$C_{1234} = x_1 x_2 x_3 x_4$,
%and
%$\Lambda = 16 \, C_{1234}$.
%
we only list the final \refeq{xr}: %characteristic equation (\ref{eq:xr}):
%after the square-free reduction:
%(with six square-free cyclic polynomials):
%
\begin{small}
\begin{align}
A_4\Rlam
%& = \left|
%\begin{array}{cccccc}
% \lambda  & 0               & 0                 & 0     & 0     & -16 \\
% 64 R     & \lambda-16(R+1) & 16 R              & 0     & -16R  & 0 \\
% -64 R^2  & 16R(R+2)        & \lambda-16(R^2+1) & -32 R & 16R   & 0 \\
% -32 R^2  & 16R(R+1)        & -16R              & \lambda-16(R^2+1)  & 0 & 0 \\
% 64 R^3   & -16R^2(R+3)     & 16R(R+2)        & 32 R (R+1)       & \lambda-16(R+1) & 0 \\
% -16R^4   & 16R^3           & - 16R^2           & -16R^2            & 16R & \lambda - 16
%\end{array}
%\right | \nonumber  \\
& \equiv
    \Big[
    4096 R^6-12288 R^5 +
    256 (\lambda + 48) (R^4 - R^3) \notag \\
& -16 (\lambda + 32)(\lambda - 16) R^2
      -(\lambda - 16)^3 \Big] \notag \\
& \quad
    \big[(16 R^2 + \lambda)^2 - 16 (2R+1)^2 \lambda \big] \notag \\
& \quad
    \big[ 16 (R-1)^2 - \lambda \big]
 = 0.
\tag{\ref{eq:xr}-4}
\label{eq:xr4}
\end{align}
\end{small}
%
%
%
The first (sextic) factor, denoted as $P_4\Rlam$,
  is responsible for the 4-cycle.
%
At the onset point, % $\lambda = 1$.
we have %it can be factorized as
$ P_4(R, 1)
= (
    5 - 4R
  )
\, \big[
  (4R + 1)^2 + 4
  \big]
\, \big[
  108 - (4R - 3)^3
  \big]$,
%
%
%
with two real roots:
$R_a' = 5/4$
%\big($r_a' = 1+\sqrt{6} \approx 3.4495$\big)
for the period-doubled 2-cycle,
  %\big[compare \reffigs{cobweb}(b) and (d)\big],
and
$R_a = (3+\sqrt[3]{108})/4$
%\big($r_a = 1+\sqrt{4+\sqrt[3]{108}} \approx 3.9601$\big)
for an original 4-cycle. %\big[\reffig{cobweb}(e)\big].
%
At the bifurcation point, % ($\lambda = -1$),
we have $P_4(R, -1)
 = 4096 R^6 - 12288 R^5 + 12032 (R^4 - R^3)
  + 8432 R^2 + 4913$.
%which upon $R \rightarrow r(r-2)/4$ yields the same polynomial
%obtained previously
%\cite{stephenson1, stephenson3, bailey2, kk1, lewis}.
%
%The only two positive roots $r_b \approx 3.9608$
%and $r_b' \approx 3.5441$ of the sextic equation
%correspond to $r_a$ and $r_a'$ respectively.
%
%Since the 4-cycle problem were tackled mainly on computers
%  in the previous studies,
%  we also supply a compact elementary derivation
%  in the Appendix.

\subsection{\label{sec:primfac}Polynomials of $n$-cycles}

@@

%\begin{theorem}
%  The polynomial $P_n\Rlam$ of all $n$-cycles
%  is a factor of the characteristic polynomial $A_n\Rlam$,
%  and can be computed as
  \begin{equation}
    P_n\Rlam
    = \prod_{cd = n} B_d^c\Rlam^{\mu(c)},
  \label{eq:primfac}
  \end{equation}
where
  $B_d^c\Rlam \equiv \prod_{k=1}^c A_{d}(R, e^{2k\pi i/c} \lambda^{1/c})$,
  particularly, $B_n^1\rlam = A_n\rlam$,
  and
  $\lambda^{1/c}$ is a complex $c$th root of $\lambda$;
  the outer product runs over all divisors $d$ of $n$.
%\end{theorem}
%
%
%\begin{remark}
%Here, ``minimal'' means that the polynomial contains all zeros.
%It happens to be irreducible for a general $\lambda$.
%At $\lambda = 1$, the polynomial of $R$, however, can be factorized (see later).
%\end{remark}
%
%\begin{remark}
The M\"obius function $\mu(n)$ is $(-1)^k$
  if $n$ is the product of $k$ distinct primes,
  or 0 if $n$ is divisible by a square of a prime;
%Thus,
$\mu(n) = 1, -1, -1, 0, -1, 1, -1, \ldots,$
  from $n = 1$
%The $\mu(n)$ is useful for inversion:
%$g(n) = \sum_{d|n} \mu(n/d) \, h(d)$
%if and only if $h(n) = \sum_{d|n} g(d)$
 \cite{hardy}.
%\end{remark}
%

%\begin{remark}
%$B_{d,c}\lambda$ is a polynomial of $\lambda$.
Despite the radical argument $\lambda^{1/c}$,
  the product
  $B_d^c\Rlam = \prod_{k=1}^c A_d(R, e^{2k\pi i/c} \lambda^{1/c})$
  is a polynomial of $\lambda$,
  for it is invariant under
  $\lambda \rightarrow e^{2\pi i} \lambda$.
Also, $\deg_\lambda B_d^c\rlam = \deg_\lambda A_d\rlam$.
For convenience, we will drop the $R$ in ``$\Rlam$'' or the arguments entirely below
  till the end the next section.
%
%\end{remark}

@@

%\subsection{\label{sec:degR}Degrees in $R$}
%
%
%
%In the numerical implementation
%  (cf. \refsec{end}),
%  the degrees in $R$ of the polynomials are needed
%  for evaluating the determinant by polynomial interpolation.
%We now show that
%\begin{equation}
%\deg_R A_n\Rlam = \textstyle\sum_{d|n} \varphi(n/d) 2^{d-1},
%\label{eq:degRA}
%\end{equation}
%%
%The first few values are
%1, 3, 6, 12, 20, 42, 70, 144, 270, 540, \dots, from $n =1$.%1034, 2112, 4108, \dots,
%
%From \refeq{xr},
%  every term in the determinant $|\vct A|$
%  can be written as
%  $\prod_{p} A_{p,q}$,
%  in which, $p$ runs through all cyclic polynomials,
%  and $\{q = q(p)\}$ is a permutation of $\{p\}$.
%Thus, $\deg_R |\vct A|$ is
%  reached by the term that maximizes the product.
%Recall that $A_{p,q}$ results from reducing
%  $\Lambda(\vx{n}) \, C_p(\vx{n})$ to $C_q(\vx{n})$
%  by a sequence of substitutions ${x_k}^2 = R-x_{k+1}$.
%It follows that to increase $\deg_R$ by 1,
%  $\deg_x$ is decreased by at least 2.
%Accordingly,
%  $\deg_R A_{p,q} \le \deg_x(\Lambda \, C_p) - \deg_x C_q$.
%Now running $p$ and $q$ through the $\NB$ cyclic polynomials
%  and using $\deg_x \Lambda(\vx{n}) = n$
%  yields
%\[
%  \deg_R |\vct A | = \deg_R \textstyle\prod_{p} A_{p,q} \le n \, \NB(n).
%\]
%%
%The equality is achieved, if each $q$ is the complement $\bar p$ of $p$,
%  that is, if $C_p$ is generated from $\prod_k x_k^{e_k}$,
%  $C_{\bar p}$ can be generated from $\prod_k x_k^{1-e_k}$
%%  (which may be the same polynomial as $C_p$).
%For example, for $n=4$,
%  $C_1 = x_1 + \cdots + x_4$
%  is the complement of
%  $C_{123} = x_2 x_3 x_4 + \cdots + x_1 x_2 x_3$,
%  while $C_{12}$ is its own complement.
%Finally, we reach \refeq{degRA} by \refeq{necklace}.
%
%
%Now, taking the degrees in $R$ of \refeq{primfac} and \refeq{origfac}
%  yields two known formulas
%\cite{mira, stephenson1, blackhurst}:
%%
%\begin{subequations}
%\begin{align*}
%\deg_R P_n\Rlam &\equiv \beta(n)
%  = \sum_{d|n} \mu\Big(\frac{n}{d}\Big) 2^{d-1}, \\
%\deg S_n(R) &= \beta(n)
%  - \sum_{d | n, d < n} \beta(d) \, \varphi\Big(\frac{n}{d}\Big).
%\end{align*}
%\label{eq:degR}
%\end{subequations}
%%
%The first few $\deg_R P_n\Rlam$ are
%  1, 1, 3, 6, 15, 27, 63, 120, 252, 495, \dots, %1023, 2010, 4095, \dots,
%and $\deg S_n(R) = $
%  1, 0, 1, 3, 11, 20, 57, 108, 240, 472, \dots, % 1013, 1959, 4083, \dots,
%from $n = 1$.
%

@@

The most time-consuming step of the algorithm in \refsec{algo}
is the evaluation of the determinant $A_n\Rlam$ in the characteristic
equation (\ref{eq:xr}).
%
When implemented on a computer,
it is helpful to use polynomial interpolation:
we evaluate $A_n\Rlam$ at $\deg_R A_n + 1$ values of $R$,
e.g., $R = 0, \pm1, \pm2,\ldots$, then piece them together
to a polynomial.
%
By using \refeq{primfac}, we only need $\deg_R P_n + 1$ values
  to directly compute $P_n\Rlam$.
%
%
Here,
$\deg_R A_n = \sum_{d|n} \varphi(n/d) 2^{d-1}$,
and $\deg_R P_n = \sum_{d|n} \mu(n/d) 2^{d-1}$
  \cite{mira, stephenson1}.
%
The strategy also allows a trivial parallelization.
